Supplement to "Properties and Bayesian fitting of restricted Boltzmann machines

### 2018-09-04 writing/paper-rbm/appendices/appendix.Rmd
more wording changes

>>>>

# Small example illustrating approximation of an arbitrary binary distribution

For simplicity, consider a model with two visible variables $(V_1,V_2)$ and one hidden $H_1$ so that there are four possible data realizations for $(V_1,V_2)$  given by $(\pm 1, \pm 1)$ and we may express the model probabilities
as
$$
P(V_1=v_1,V_2=v_2| \theta_{v_1}, \theta_{v_2}, \theta_{h_1}, \theta_{11},\theta_{21}) \propto  \exp\left( v_1  \theta_{v_1} + v_2 \theta_{v_2}\right) \sum_{h \in \{\pm 1\}}\exp\left( h[ \theta_{h_1}  + \theta_{11} v_1 + \theta_{21} v_2] \right),
$$
for $(v_1,v_2)\in \{-1, 1\}^2$, in terms of real-valued parameters $\theta_{v_1}, \theta_{v_2}, \theta_{h_1}, \theta_{11},\theta_{21}$. Given some specified cell probabilities, say
\begin{equation}
(\#eq:0)
0\leq p_{(-1,-1)},\;p_{(1,-1)}, \;p_{(-1,1)},\;p_{(1,1)},
\end{equation} 
for the outcomes $(\pm 1, \pm 1)$, the online appendix provides technical details to show parameter values $(\theta_{v_1}, \theta_{v_2}, \theta_{h_1}, \theta_{11},\theta_{21})$ may be chosen to match any given cell probabilities with arbitrary closeness.  In fact, when the cell probabilities \@ref(eq:0) are all strictly positive, parameters in the RBM model can be specified to reflect these probabilities exactly. And, when one or more of the cell probabilities \@ref(eq:0) are zero, the corresponding RBM probabilities $P(V_1=v_1,V_2=_v2| \theta_{v_1}, \theta_{v_2}, \theta_{h_1}, \theta_{11},\theta_{21})$ may never be identically zero (due to exponential terms in the model) but parameters can be still selected to make the appropriate RBM cell probabilities arbitrarily small.

For illustration, considering the simple example from above with $n$ iid observations, each consisting of two realized visibles $(V_1,V_2)$ the empirical cell frequencies from the sample are the highest likelihood choices for the cell probabilities $p_{(-1,-1)},p_{(1,-1)}, p_{(-1,1)},p_{(1,1)}$ in \@ref(eq:0) before any parametric model is introduced for improved analysis, and the discussion above indicates that RBM model parameters can be chosen to re-create this empirical distribution to an arbitrarily close degree.
